Displacement of fishing effort by Large Scale Marine Protected Areas[^*]

### 2018-10-21 rmd/Manuscript.Rmd
Updated all

---
title: "Displacement of fishing effort by Large Scale Marine Protected Areas[^*]"
subtitle: "Updated on `r Sys.Date()`"
author: |
  Juan Carlos Villaseñor-Derbez^1^
  John Lynham^2^
abstract: "Large-scale Marine Protected Areas (LSMPAs) have seen a significant increase over the last years. Fishing effort is effectively eliminated within these protected areas upon implementation. The benefits of reducing effort have been largely studied, and include increases in abundance, biomass, and diversity within the bounded regions. These no-take zones may produce spillover effects, which provide fish for outside areas. However, the economic and ecological implications of displacing fishing effort are not yet fully understood. Novel data products that track fishing effort at the vessel-level allow us to identify changes in fleet- and vessel-level behavior upon the implementation of protected areas, as well as how these redistribute. This papers evaluates the implications of implementing LSMPA, by evaluating changes in fishing hours, showing that vessels in the effected region reduce fishing effort after the implementation of PIPA. Our results are robust to a set of specifications. We also track the relative spatial allocation of fishing events thorugh time, and identify that areas closer to PIPA show an increase in relative fishing hourse due to the displacement of PIPA-fishing vessels. Our results not only provide an impact evaluation of the effect of LSMPAs on fishing activity, but provide insights into vessel redistribution dynamics, which may have ecological and economic implications."
date: "
^1^Bren School of Environmental Science and Management, University of California Santa Barbara, Santa Barbara, CA\\newline
^2^Department of Economics, University of Hawaii at Manoa, Honolulu, HI"
output:
  bookdown::pdf_document2:
    fig_height: 3.5
    fig_width: 6
    toc: no
    citation_package: natbib
    keep_tex: yes
    number_sections: true
fontsize: 11pt
header-includes:
  - \usepackage{float}
  - \floatplacement{figure}{H}
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage[bottom]{footmisc}
  - \usepackage{pdflscape}
  - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
bibliography: references.bib
biblio-style: chicago
---

@@

```{r, echo = F, message = F, warning = F}
suppressPackageStartupMessages({
  library(startR)
  library(sf)
  library(tidyverse)
  library(magrittr)
})

@@

The cutoff at which an MPA is considered to be a LSMPA ranges from areas larger than 30,000 km^2^ as defined by \citet{desanto_2013} or areas larger than 250,000 km^2^, as defined by \citep{toonen_2013}. Figure \ref{fig:LSMPAs_map} shows LSMPAs that meet the later, and that are fully no take. LSMPAs are often implemented in the pelagic environment, where the dominant human activity is industrial fishing \citep{gray_2017,kroodsma_2018}. The early literature on LSMPAs focused on the inherent challenges and difficulties that come with a pelagic environment. \citet{kaplan_2010} claimed that very large MPAs would result in excessive opportunity costs and that these would be difficult to enforce. \citet{game_2009} suggested that most of the challenges could be overcome with the incorporation of technology, in what then became known as Dynamic Ocean Management \citep{maxwell_2015}. The effectiveness and difficulties of implementating LSMPAs are not the focus of this paper \footnote{But see \citet{singleton_2014}, who provide an objective discussion of pros and cons of LSMPAs.}. Instead, we focus on the effect and implications of the ones that *have been* implemented.

@@

LSMPAs were erroneously assumed to have little social implications due to their remoteness. However, there have been calls to incorporate the human dimensions into LSMPAs management and evaluation \citep{agardy_2011,gray_2017}. Most research incorporating these dimensions has focused on governance and enforcement of LSMPAs (*i.e.* \citet{alger_2017,christie_2017}), but they are yet to be the focus of economic analyses \citep{gray_2017}. Overall, there has been little empirical work regarding LSMPAs. Recent technological advances in vessel-detection systems allows for the discovery and advancement of many important facets of LSMPAs. For example, \citep{mcdermott_2018} show that the anticipation of a LSMPA can lead to preemptive overfishing, which can erode or delay the expected benefits of the intervention. \citet{white_2017} combine shark tags and vessel-tracking data to demonstrate that the fairly large Palmyra Atoll National Wildlife Refuge (54,000 Km^2^) protectes two thrids of the tagged grey reef sharks by effectively excluding fishing effort. More recently, \citep{bradley_2018} use similar data to highlight cases of potential illegal shark fishing *inside* a 2 million km^2^ shark sanctuary. To date, no studies have evaluated the displacement of fishing effort due to LSMPA implementation.

@@

The empirical research that has been done in custumary sized MPAs suggest that resource users may show idiosyncratic responses. For example, \citet{stevenson_2013} show that a network of MPAs displaced fishing effort farther away from ports, resulting in higher *perceived* costs, and increases in catch per unit effort. \citet{cabral_2017} analyse the redistribution of fishing and non-fishing vessels following the implementation of a network of MPAs in California, and find that commercial dive boats follow a fishing-the-line pattern, while some fishing boats follow an ideal free distribution. More recently \citet{elahi_2018} used satellite tracking data to show that a temporal spatial closure caused trawlers to maintain effort but apply it more intensively elsewhere, particularly along the borders and closer to shore. The way in which fishers react to a spatial closure can have major implications in its outcome \citep{smith_2003,hilborn_2006} highlighting the need to understand how fishers react to the implementation of LSMPAs, and fishing effort changes and is spatially redistributed. However, all these closures took place within Exlcusive Economic Zones, where other regulations exist. This may not always be the case for LSMPAs, where often the entire EEZ is converted into a LSMPA, leaving fishers with the options of moving to the high seas or other countries' EEZs.

@@

The cooperation that emerged thanks to the PNA allowed for subsequent agreements that strengthened fisheries management, like the Palau Agreement, which limited the number of purse seiners at 205 vessels from 1995-2007[^a]. However, the most notable regulation is their approach to manage fishing effort: a Vessel Day Scheme (VDS) implemented in 2007 \citep{havice_2013}. This effectively modified how fishing effort was managed, from number of vessels under the Palau Agreement to fishing hours. The VDS works as follows: Each year, scientific advisors recomend a total number of fishing vessel-days per year. Hours are allocated to each PNA country based on catch history, and they then use or sell fishing rights to other non-PNA countries \citep{aqorau_2018}. While the effectiveness of this scheme has been debated in terms of meeting their fishery management and conservation objectives, the licensing significantly contributes to the economy of these island nations \citep{havice_2010}.

@@

```{r, fig.cap = "\\label{fig:PNA_map}Map of the Exclusive Economic Zones (EEZs) of the region of interest. A solid line indicates countries that belong to the PNA, while a dashed line indicates all others. A red line indicates the Kiribati EEZ, and a solid red polygon delineates PIPA."}
countries <- c("PIPA", "KIR", "HS", "ASM", "COK", "FSM", "MHL", "NRU", "PNG", "SLB", "TKL", "TUV", "UMI"< "FJI")

@@

eez <- read_sf(dsn = here::here("data", "spatial", "EEZ_subset"),
               layer = "EEZ_subset") %>% 
  filter(ISO_Ter1 %in% countries) %>% 
  mutate(KIR = ISO_Ter1 == "KIR") %>% 
  group_by(ISO_Ter1, PNA, KIR) %>% 
  summarize()

ggplot() +
  geom_sf(data = eez, aes(color = KIR, linetype = !PNA), fill = "transparent") +
  geom_sf(data = pipa, fill = "red") +
  geom_sf(data = small_coast, size = 1, color = "black") +
  ggtheme_plot() +
  scale_color_manual(values = c("black", "red")) +
  theme(legend.position = "none")
```

The main tuna species caught in the region are skipjack (*Katsuwonus pelamis*), yellowfin (*Thunnus albacares*), albacore (*Thunnus alalunga*) and bigeye (*Thunnus obesus*). From these, the first two are amongst the top-10 species represented in global fisheries production statistics, with 2016 catches increasing relative to the 2005-2014 average \citep{fao_2018}. Theis region of the Pacific has historically accounted for a large portion of tuna catches \citep{aqorau_1997}. Today, the PNA controls close to 50% of the global skipjack tuna production \citep{pna_website_2018}. A large portion of these catches derive from purse seine vessels licensed under the VDS. Fishing vessels from Australia, New Zeland, China, France, Korea, Japan, the Philippines, Taiwan, and the United States participate in the purse-seining VDS.

@@

Our analysis focuses on purse seine vessels, the most important fishery for PNA countries\footnote{We perform some of the same analyses for longliners and include them in the Appendix}. We identify a total of 103 purse seiners that fished in PNA waters at least once before 2015\footnote{New vessels that entered the fishery after 2015 were not exposed to the policy intervention in the pre-treatment period and are therefore exlcuded from our analyses.}. These vessels represent over 26 million individual observations for the 2012 - 2017 period. We identify 65 vessels that have fished inside PIPA at least once since 2012. From these, 62 did so before the anouncement (*i.e.* 09/01/2014 *sensu* \citep{mcdermott_2018}) but are only able to track 61 for the complete periof of study before and after its implementation\footnote{1 vessel never fished again after August 18, 2013.}. On the other hand, 38 vessels never fished inside PIPA but we only have data for 26 vessels before and after.

Therefore, our treatment group contains all purse seiners (n = 61) that fished within PIPA at least once before the anouncement, and that continued to fish elsewhere after the January 2015 implementation. Vessels in the control group meet the following two conditions: i) vessels never fished within PIPA waters, and ii) vessels have fished in surrounding areas (*i.e.* PNA-countries' EEZ) before and after PIPA closure (n = 26).

We include three additional definitions of control groups as a robustness check: one with only vessels that belong to PNA countries (n = 7), and one that excludes Chinese vessels (n = 21). Our third control is made up of Japanese purse seiners that fish in the Pacific but that never fished inside PIPA (n = 27)\footnote{Using Taiwanese vessels was the original idea, but there are only 4 vessels with pre-2015 data. I need to further identify propper Taiwanese vessels, as I am currently using the entire Pacific ocean}. Our definition of treatment and control groups leaves us with 61 and 21 treated and control vessels, which have just over 22 million observations where about 22% are identified as fishing events.

For each vessel we calculate total daily fishing hours and obtain pannel data with 37,800 observations. Table \ref{tab:baci_n_s} shows the number of vessels following a BACI design, as well as the fishing hours, before and after PIPA. Fig. \ref{fig:all_vessels} shows that mean fishing hours for purse seiners have an abrupt increase, just before January 1st, 2015. This trend is observed for both treated and control groups. Across all measures, the treatment and control vessels follow similar patterns, confirming our claim that the control group provides a plausible counterfactual. Figure \ref{fig:baci_strict} provides a visual representation of the vessel-level fishing events that make up each group through time. We use this data to answer our key questions: How do the 61 purse seiners modify their behavior as compared to the different control groups, and where do they go after the spatial closure? The following section describes our empirical identification strategies.

@@

$$
y_{i,t} = \alpha + \beta_1 Post_t + \beta_2 Treat_i + \beta_3 Post_t \times Treat_i + \phi_t + \gamma_i + \epsilon_{i,t}
$$

Where $y_{i,t}$ is the variable of interest for vessel $i$ in time period $t$. A dummy variable $Post_t$ takes the value of 0 for all dates prior to PIPA implementation and a value of 1 for all dates including and following PIPA implementation. $Treat_i$ is a dummy variable indicating whether a vessel belongs to the control ($Treat_i = 0$) or treatment ($Treat_i = 1$) group. $\alpha$ is the standard intercept, $\beta_1$ captures the temporal trend change, $\beta_2$ captures the difference between treated and control groups, and $\beta_3$ is our parameter of interest: de DiD estimate capturing the treatment effect. Finally, $\phi_t$ and $\gamma_i$ represent month-, and flag-level dummies that account for seasonality or country-level management interventions\footnote{We test for other more complex specifications that interact a quarterly dummy or year-month dummies with the treatment group and find qualitatively the same results.}.

@@

Our second part of the analyses focuses on the redistribution of fishing effort. The role of institutions may play an important role. As stated before, LSMPAs often span the entirety of an EEZ. In the case of purse seining in the PNA, vessels purchase access to the fishery at the beginning of the season. If a vessel decides to fish within the PNA, they've already made the decission to pay for access, expecting higher returns from fishing there than in the High Seas. In the particular case of PIPA, one would expect that a vessel holding a permit to fish in I-Kiribati waters would then move to other fishing grounds within Kiribati after the closure. Alternatively, if a vessel fishes illegaly in I-Kiribati waters before the implementation, one would expect them to reallocate to other regions with equal or less enforcement. In this case, they might then choose to move to other I-Kiribati waters, waters of countries that have similar enforcement levels, or the High Seas.

We discretize spatial units by using a polygon for PIPA \footnote{we would expect to see a decrease here} and distinct spatial units for each EEZ of each country. Some vessels might shift from EEZs into the high seas, but we are interested in knowing *where* in the high seas, so we incorporate additional regions by using a 1 degree buffer of the high seas arround each of the EEZ regions. The rest of the high seas are merged into a single spatial unit. For example, if we were to do this only for Kiribati, we would have 8 spatial units: PIPA, three EEZs, three 1-degree buffers of high seas arround each EEZs, and the rest of the high seas. Whenever the buffers overlapped between themselves, we randocmply clipped one on to the other. EEZs that had sporadic fishing events were pooled into a group of "others".

To evaluate this change in effort allocation, we regress our variable of interest (*i.e.* fishing hours) on the interaction between a dummy variable indicating the policy intervention and a dummy variable for countries. This gives us the by-country change in proportional allocation of fishing effort:

$$
y_{i,t} = \alpha + \beta_1Post + \beta_{2,i}Country + \beta_{3,i}Post_t \times Country_i + \epsilon_{i,t}
$$

Our variable of interest, $y_{i,t}$ represents the proportion of fishing hours that country $i$ receives at time $t$. $Post$ also represents a policy dummy that takes the value of 0 for all dates before implementation of PIPA, and 1 otherwise. $Country$ is a dummy variable for countries for the spatial units defined above. Our parameter of interest is $\beta_{1,i}$, which captures the country-level change in proportional fishing effort.

<!-- Like before, we include a dummy variable for months and allow for a second degree polynomial trend through time. -->

@@

```{r}
ba_disp <- displacement_data %>% 
  filter(gear == "purse_seines") %>% 
  group_by(country, post) %>% 
  summarize(h_prop_m = mean(h_prop)) %>% 
  spread(post, h_prop_m) %>% 
  mutate(change = (`TRUE` - `FALSE`) * 100) %>% 
  select(country, change)

knitr::kable(x = ba_disp,
             digits = 2,
             booktabs = T,
             caption = "\\label{tab:ba_disp}Changes in the relative allocation of fishing effort by region (EEZ, PIPA, high seas) and gear.") %>%
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

@@

Our DiD analysis shows an overall increase in purse seine fishing hours, even after accounting for the introduction of new satellites (Table \ref{tab:did}). This coefficient estimate is consistent for different model specifications and across groups of treatment and controls. These effectively represent the patterns observed in Figure \ref{fig:all_vessels}. The $\beta_3$ coefficient indicating our treatment effect suggests that, relative to the control, treated vessels fish less, in the order of 0.5 hours per day. Another way to interpret this is that the increase in fishing effort by treated vessels has ocurred at a lower rate than control vessels. This result is robust and significant ($p < 0.05$) for all model specifications using our main control group, with the effect varying between $\beta_3 = -0.515$ and $\beta_3 = -0.934$. These values are equivalent to a 15 - 28 reduction in fishing hours per month.

Regressions coefficients for our DiD analysis are shown in Table \ref{tab:did}. Columns 1 -3 use all data for controls. Columns 4 - 6 use only PNA vessels as controls, and columns 7 - 9 exclude Chinese vessels. Columns 10 and 11 use Japanese vessels in the Pacific as controls, and hence don't include flag FEs to avoid perfect colinearity between Japanese flag and control\footnote{Results of the same analysis is shown for longliners in \ref{tab:long}}. It must be noted, however, that when reducing the linear structure of the $Pre \times Post$ design and instead interacting the treatment dummy with quarterly or year-month combinations we are not able to reject the null hypotheses of no change (Figures \ref{fig:q1}-\ref{fig:ym4}).

```{r, echo = F, fig.cap = "\\label{fig:all_vessels}Daily fishing hours for all vessels in our main treatment-control groups. Solid straight lines show a linear trend by period (pre-psot) and treatment. The other red and blue lines show monthly avarages. Vertical dashed line indicates PIPA closure."}
effort_by_vessel %>% 
  filter(baci_strict) %>% 
  mutate(date = lubridate::date(paste(year, month, 15, sep = "/"))) %>%
  ggplot(aes(x = date, y = hours, color = treated)) +
  geom_jitter(shape = ".", height = 0, alpha = 0.1) +
  stat_summary(geom = "line", fun.y = mean, size = 1) +
  geom_smooth(method = "lm", aes(group = paste(treated, post))) +
  ggtheme_plot() +
  scale_color_brewer(palette = "Set1") +
  geom_vline(xintercept = lubridate::date(paste(2015, 01, 01, sep = "/")), linetype = "dashed")
```

>>>>

---
title: "Displacement of fishing effort by Large Scale Marine Protected Areas[^*]"
subtitle: "Updated on `r Sys.Date()`"
author: |
  Juan Carlos Villaseñor-Derbez^1^
  John Lynham^2^
abstract: "Large-scale Marine Protected Areas (LSMPAs) have seen a significant increase over the last years. Fishing effort is effectively eliminated within these protected areas upon implementation. The benefits of reducing effort have been largely studied, but little empirical works evaluate what happens to the effort after an MPA is created. The economic and ecological implications of displacing fishing effort are not yet fully understood. We use identification of fishing activity via Automatic Identification Systems (AIS) and causal inference techniques to provide the first analyses of behavioral changes and spatial redistribution of tuna purse seiners due to the implementation of a Large Scale Marine Protected Area in the Pacific Ocean. Our work provides three main findings: 1) aggregate fishing effort remains relatively unaffected; 2) Vessels that fished inside the protected area redistribute to adjacent waters; and 3) We observe a crowding effect for the first years after implementation. Our results not only provide an impact evaluation of the effect of LSMPAs on fishing activity, but provide insights into vessel redistribution dynamics, which may have ecological and economic implications. As countries continue to implement LSMPAs as a way to reach the stated 10% target of ocean protection, managers should consider how fishing effort will change in space and through time to ensure that fishing effort is not just displaced elsewhere, leading to overfishing in adjacent waters. While LSMPAs can provide a wide range of benefits, their implementation must be accompained with traditional fisheries management to maximize their effectiveness."
date: "
^1^Bren School of Environmental Science and Management, University of California Santa Barbara, Santa Barbara, CA\\newline
^2^Department of Economics, University of Hawaii at Manoa, Honolulu, HI"
output:
  bookdown::pdf_document2:
    fig_height: 3.5
    fig_width: 6
    toc: no
    citation_package: natbib
    keep_tex: yes
    number_sections: true
fontsize: 11pt
header-includes:
  - \usepackage{float}
  - \floatplacement{figure}{H}
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage[bottom]{footmisc}
  - \usepackage{pdflscape}
  - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
# bibliography: references.bib
biblio-style: chicago
---

@@

```{r, echo = F, message = F, warning = F}
suppressPackageStartupMessages({
  library(raster)
  library(startR)
  library(sf)
  library(tidyverse)
  library(magrittr)
})

@@

The cutoff at which an MPA is considered to be a LSMPA ranges from areas larger than 30,000 km^2^ as defined by \citet{desanto_2013} or areas larger than 250,000 km^2^, as defined by \citep{toonen_2013}. Figure \ref{fig:LSMPAs_map} shows LSMPAs that meet the later, and that are fully no take. LSMPAs are often implemented in the pelagic environment, where the dominant human activity is industrial fishing \citep{gray_2017,kroodsma_2018}. The early literature on LSMPAs focused on the inherent challenges and difficulties that come with a pelagic environment. \citet{kaplan_2010} claimed that very large MPAs would result in excessive opportunity costs and that these would be difficult to enforce. \citet{game_2009} suggested that most of the challenges could be overcome with the incorporation of technology, in what then became known as Dynamic Ocean Management \citep{maxwell_2015}. The effectiveness and difficulties of implementing LSMPAs are not the focus of this paper \footnote{But see \citet{singleton_2014}, who provide an objective discussion of pros and cons of LSMPAs.}. Instead, we focus on the effect and implications of the ones that *have been* implemented.

@@

LSMPAs were erroneously assumed to have little social implications due to their remoteness. However, there have been calls to incorporate the human dimensions into LSMPAs management and evaluation \citep{agardy_2011,gray_2017}. Most research incorporating these dimensions has focused on governance and enforcement of LSMPAs (*i.e.* \citet{alger_2017,christie_2017}), but they are yet to be the focus of economic analyses \citep{gray_2017}. Overall, there has been little empirical work regarding LSMPAs. Recent technological advances in vessel-detection systems allows for the discovery and advancement of many important facets of LSMPAs. For example, \citep{mcdermott_2018} show that the anticipation of a LSMPA can lead to preemptive overfishing, which can erode or delay the expected benefits of the intervention. \citet{white_2017} combine shark tags and vessel-tracking data to demonstrate that the fairly large Palmyra Atoll National Wildlife Refuge (54,000 Km^2^) protects two thirds of the tagged grey reef sharks by effectively excluding fishing effort. More recently, \citep{bradley_2018} use similar data to highlight cases of potential illegal shark fishing *inside* a 2 million km^2^ shark sanctuary. To date, no studies have evaluated the displacement of fishing effort due to LSMPA implementation.

@@

The empirical research that has been done in customary sized MPAs suggest that resource users may show idiosyncratic responses. For example, \citet{stevenson_2013} show that a network of MPAs displaced fishing effort farther away from ports, resulting in higher *perceived* costs, and increases in catch per unit effort. \citet{cabral_2017} analyse the redistribution of fishing and non-fishing vessels following the implementation of a network of MPAs in California, and find that commercial dive boats follow a fishing-the-line pattern, while some fishing boats follow an ideal free distribution. More recently \citet{elahi_2018} used satellite tracking data to show that a temporal spatial closure caused trawlers to maintain effort but apply it more intensively elsewhere, particularly along the borders and closer to shore. The way in which fishers react to a spatial closure can have major implications in its outcome \citep{smith_2003,hilborn_2006} highlighting the need to understand how fishers react to the implementation of LSMPAs, and fishing effort changes and is spatially redistributed. However, all these closures took place within Exclusive Economic Zones, where other regulations exist. This may not always be the case for LSMPAs, where often the entire EEZ is converted into a LSMPA, leaving fishers with the options of moving to the high seas or other countries' EEZs.

@@

The cooperation that emerged thanks to the PNA allowed for subsequent agreements that strengthened fisheries management, like the Palau Agreement, which limited the number of purse seiners at 205 vessels from 1995-2007[^a]. However, the most notable regulation is their approach to manage fishing effort: a Vessel Day Scheme (VDS) implemented in 2007 \citep{havice_2013}. This effectively modified how fishing effort was managed, from number of vessels under the Palau Agreement to fishing hours. The VDS works as follows: Each year, scientific advisers recommend a total number of fishing vessel-days per year. Hours are allocated to each PNA country based on catch history, and they then use or sell fishing rights to other non-PNA countries \citep{aqorau_2018}. While the effectiveness of this scheme has been debated in terms of meeting their fishery management and conservation objectives, the licensing significantly contributes to the economy of these island nations \citep{havice_2010}.

@@

```{r, fig.cap = "\\label{fig:PNA_map}Map of the Exclusive Economic Zones (EEZs) of the region of interest. Countries that belong to the PNA are shown in blue, while empty polygons indicates all others. A red line indicates the Kiribati EEZ, and a solid red polygon delineates PIPA."}
countries <- c("PIPA", "KIR", "HS", "ASM", "COK", "FSM", "MHL", "NRU", "PNG", "SLB", "TKL", "TUV", "UMI", "FJI", "NIU", "TON", "WSM", "WLF", "VUT", "NCL")

@@

eez <- read_sf(dsn = here::here("data", "spatial", "EEZ_subset"),
               layer = "EEZ_subset") %>% 
  filter(ISO_Ter1 %in% countries) %>% 
  mutate(KIR = ISO_Ter1 == "KIR",
         PNA = ifelse(PNA, "PNA", "Non-PNA")) %>% 
  group_by(ISO_Ter1, PNA, KIR) %>% 
  summarize()

ggplot() +
  geom_sf(data = eez, aes(color = KIR, fill = PNA), alpha = 0.5) +
  geom_sf(data = pipa, fill = "red") +
  geom_sf(data = small_coast, size = 1, color = "black") +
  ggtheme_plot() +
  scale_color_manual(values = c("black", "red")) +
  scale_fill_manual(values = c("transparent", "steelblue")) +
  theme(legend.position = "none")
```

The main tuna species caught in the region are skipjack (*Katsuwonus pelamis*), yellowfin (*Thunnus albacares*), albacore (*Thunnus alalunga*) and bigeye (*Thunnus obesus*). From these, the first two are amongst the top-10 species represented in global fisheries production statistics, with 2016 catches increasing relative to the 2005-2014 average \citep{fao_2018}. This region of the Pacific has historically accounted for a large portion of tuna catches \citep{aqorau_1997}. Today, the PNA controls close to 50% of the global skipjack tuna production \citep{pna_website_2018}. A large portion of these catches derive from purse seine vessels licensed under the VDS. Fishing vessels from Australia, New Zealand, China, France, Korea, Japan, the Philippines, Taiwan, and the United States participate in the purse-seining VDS.

@@

Our analysis focuses on purse seine vessels, the most important fishery for PNA countries\footnote{We perform some of the same analyses for longliners and include them in the Appendix}. We identify a total of 103 purse seiners that fished in PNA waters at least once before 2015\footnote{New vessels that entered the fishery after 2015 were not exposed to the policy intervention in the pre-treatment period and are therefore excluded from our analyses.}. These vessels represent over 26 million individual observations for the 2012 - 2017 period. We identify 65 vessels that have fished inside PIPA at least once since 2012. From these, 62 did so before the announcement (*i.e.* 09/01/2014 *sensu* \citep{mcdermott_2018}) but are only able to track 61 for the complete period of study before and after its implementation\footnote{1 vessel never fished again after August 18, 2013.}. On the other hand, 38 vessels never fished inside PIPA but we only have data for 26 vessels before and after.

Therefore, our treatment group contains all purse seiners (n = 61) that fished within PIPA at least once before the announcement, and that continued to fish elsewhere after the January 2015 implementation. Vessels in the control group meet the following two conditions: i) vessels never fished within PIPA waters, and ii) vessels have fished in surrounding areas (*i.e.* PNA-countries' EEZ) before and after PIPA closure (n = 26).

We include three additional definitions of control groups as a robustness check: one with only vessels that belong to PNA countries (n = 7), and one that excludes Chinese vessels (n = 21). Our third control is made up of Japanese purse seiners that fish in the Pacific but that never fished inside PIPA (n = 27)\footnote{Using Taiwanese vessels was the original idea, but there are only 4 vessels with pre-2015 data. I need to further identify proper Taiwanese vessels, as I am currently using the entire Pacific ocean}. Our definition of treatment and control groups leaves us with 61 and 21 treated and control vessels, which have just over 22 million observations where about 22% are identified as fishing events.

For each vessel we calculate total daily fishing hours and obtain panel data with 37,800 observations. Table \ref{tab:baci_n_s} shows the number of vessels following a BACI design, as well as the fishing hours, before and after PIPA. Fig. \ref{fig:all_vessels} shows that mean fishing hours for purse seiners have an abrupt increase, just before January 1st, 2015. This trend is observed for both treated and control groups. Across all measures, the treatment and control vessels follow similar patterns, confirming our claim that the control group provides a plausible counterfactual. Figure \ref{fig:baci_strict} provides a visual representation of the vessel-level fishing events that make up each group through time. We use this data to answer our key questions: How do the 61 purse seiners modify their behavior as compared to the different control groups, and where do they go after the spatial closure? The following section describes our empirical identification strategies.

@@

$$
y_{i,t} = \alpha + \beta_1 Post_t + \beta_2 Treat_i + \beta_3 Post_t \times Treat_i + \phi_t + \gamma_i + \theta_t + \epsilon_{i,t}
$$

Where $y_{i,t}$ is the variable of interest for vessel $i$ in time period $t$. A dummy variable $Post_t$ takes the value of 0 for all dates prior to PIPA implementation and a value of 1 for all dates including and following PIPA implementation. $Treat_i$ is a dummy variable indicating whether a vessel belongs to the control ($Treat_i = 0$) or treatment ($Treat_i = 1$) group. $\alpha$ is the standard intercept, $\beta_1$ captures the temporal trend change, $\beta_2$ captures the difference between treated and control groups, and $\beta_3$ is our parameter of interest: de DiD estimate capturing the treatment effect. Finally, $\phi_t$ and $\gamma_i$ represent month- and flag-level dummies that account for seasonality or country-level management interventions. $\theta_t$ are dummies to account for the addition of more satellites. \footnote{We test for other more complex specifications that interact a quarterly dummy or year-month dummies with the treatment group and find qualitatively the same results.}.

@@

Our second part of the analyses focuses on the redistribution of fishing effort. The role of institutions may play an important role. As stated before, LSMPAs often span the entirety of an EEZ. In the case of purse seining in the PNA, vessels purchase access to the fishery at the beginning of the season. If a vessel decides to fish within the PNA, they've already made the decision to pay for access, expecting higher returns from fishing there than in the High Seas. In the particular case of PIPA, one would expect that a vessel holding a permit to fish in I-Kiribati waters would then move to other fishing grounds within Kiribati after the closure. Alternatively, if a vessel fishes illegally in I-Kiribati waters before the implementation, one would expect them to reallocate to other regions with equal or less enforcement. In this case, they might then choose to move to other I-Kiribati waters, waters of countries that have similar enforcement levels, or the High Seas.

We discretize spatial units by using a polygon for PIPA\footnote{we would expect to see a decrease here} and distinct spatial units for each EEZ of each country. Some vessels might shift from EEZs into the high seas, but we are interested in knowing *where* in the high seas, so we incorporate additional regions by using a 1 degree buffer of the high seas around each of the EEZ regions. The rest of the high seas are merged into a single spatial unit. For example, if we were to do this only for Kiribati, we would have 8 spatial units: PIPA, three EEZs, three 1-degree buffers of high seas around each EEZs, and the rest of the high seas. Whenever the buffers overlapped between themselves, we randomly clipped one on to the other.

We then take this spatial polygons and rasterize them to a 1-degree grid (See Figure \ref{fig:raster_rgn}), which we then use to rasterize points of fishing activity. For each cell, we calculate the total monthly fishing hours by treated and control vessels (See Figure \ref{fig:fishing_raster}). Since we are interested in the redistribution of fishing effort, we use this gridded data to  calculate the proportion of fishing hours that each region receives each month relative to the total fishing hours observed across all regions and all cells in that month.

We use this measure to evaluate the change in effort allocation by regressing it on the interaction between a year dummy and a dummy for regions:

$$
y_{i,t} = \alpha + \beta_1Year_t + \beta_{2,i}Region_i + \beta_{3,i}Year_t \times Region_i+ \epsilon_{i,t}
$$

Our variable of interest, $y_{i,t}$ represents the proportion of fishing hours that region $i$ receives at time $t$.  Years are modelled as factors ($Year_t$), using 2012 as the reference level. $Region$ is a dummy variable for regions defined above. Our parameter of interest is $\beta_{3,i}$, which captures the yearly by-country change in proportional allocation of fishing effort relative to 2012.

@@

```{r}
# Monthly  for analyses
raster_month_region <- readRDS(file = here::here("data", "monthly_rasterized_effort_by_region.rds"))

# Yearly for visualization purposes
fishing_raster <- readRDS(file = here::here("data", "rasterized_effort_by_region.rds"))

redistribution_model_data <-
  expand.grid(id = unique(raster_month_region$id),
              year = 2012:2017,
              month = 1:12) %>% 
  arrange(id, year, month) %>%
  as.tibble() %>%
  left_join(raster_month_region, by = c("id", "year", "month")) %>% 
  group_by(year, month) %>% 
  mutate(total_hours = sum(hours, na.rm = t)) %>% 
  ungroup() %>% 
  group_by(year, month, id, total_hours) %>%
  summarize(h = sum(hours, na.rm = T)) %>% 
  ungroup() %>% 
  mutate(h_prop = h / total_hours) %>% 
  mutate(h_prop = ifelse(is.na(h_prop), 0, h_prop)) %>% 
  mutate(date = lubridate::date(paste(year, month, 1, sep = "/")),
         post = year >= 2015,
         sate = case_when(date < lubridate::date("2014/06/01") ~ "1",
                          date > lubridate::date("2015/12/31") ~ "3"),
         year_c = as.factor(year))

model <- lm(h_prop ~ year_c * id, data = redistribution_model_data)
```

@@

Regressions coefficients for our DiD analysis are shown in Table \ref{tab:did}. Columns 1 -3 use all data for controls. Columns 4 - 6 use only PNA vessels as controls, and columns 7 - 9 exclude Chinese vessels. Columns 10 and 11 use Japanese vessels in the Pacific as controls, and hence don't include flag fixed-effects to avoid perfect colinearity between Japanese flag and control\footnote{Results of the same analysis is shown for longliners in \ref{tab:long}}. Our DiD analysis shows an overall increase in purse seine fishing hours, even after accounting for the introduction of new satellites (Table \ref{tab:did}). This coefficient estimate is consistent for different model specifications and across groups of treatment and controls, and effectively represents the patterns observed in Figure \ref{fig:all_vessels}.

The $\beta_3$ coefficient indicating our treatment effect suggests that, relative to the control, treated vessels fish less, in the order of 0.5 hours per day. Another way to interpret this is that the increase in fishing effort by treated vessels has occurred at a lower rate than control vessels. The effect varying between $\beta_3 = -0.515$ and $\beta_3 = -0.934$. These values are equivalent to a 15 - 28 reduction in fishing hours per month. However, this result is only robust and significant ($p < 0.01$) in the simplest specification of the Diff-in-Diff. When reducing the linear structure of the $Pre \times Post$ design and instead interacting the treatment dummy with quarterly or year-month combinations we are not able to reject the null hypotheses of no change, suggesting that the total fishing hours remains similar for the displaced vessels (Figs. \ref{fig:q1}-\ref{fig:ym4}). Since vessels don't seem to be fishing less (or at least not much less), an obvious question arises: where are they fishing now?

Figure \ref{fig:fishing_raster} provides a spatial representation of these changes. Note that 2013 effort distribution is roughly similar across vessels. Then in 2014 there is a sharp increase in fishing hours by treated vessels inside PIPA (blue paradox). In 2015 treated vessels then allocate more effort to the easternmost Kiribati EEZ (see Figure \label{fig:PNA_map}) and the high seas. In 2016 spatial distribution of fishing effort is again similar across groups. It is evident that the increase in relative fishing effort is greater for for regions closer to PIPA.

Along a similar line, Figure \ref{fig:redist_trend} shows a detailed temporal evolution of the relative allocation of fishing effort\footnote{Recall that to evaluate the redistribution of fishing effort we only track fishing vessels that belong to the treated group.}. It shows shows the change in fishing effort inside PIPA, including the preemptive fishing and immediate reduction previously reported \citep{mcdermott_2018}. I-Kiribati waters in KIR 1 and KIR 2 show an increase in fishing effort for 2015; the same pattern is observed for their respective High Seas buffers, and the general High Seas. Other regions farther away from PIPA don't show a clear pattern under visual inspection.
